Causes and effects in Dichotomous Comparative Judgments: an information-theoretical system of plausible mechanism

### 2024-08-06 index.qmd
section:background,flow; process:structure

<!-- 3. What are the advantages of DAGs? -->
DAGs offer two key advantages for modeling causality. Firstly, DAGs represent causal relations in a nonparametric and fully interactive manner. This feature implies that feasible causal analysis strategies usually do not require specifying the type of data or the nature of the functional dependence among variables [@Morgan_et_al_2014]. Secondly, regardless of complexity, DAGs can represent various causal structures using only five fundamental building blocks [@Neal_2020; @McElreath_2020]. These building blocks can be depicted in three ways: the magnified representation, the standard representation, and the structural causal model form [SCM, @Morgan_et_al_2014].

@@

<!-- 3. But what are those mysterious functions?, and how do they become statistical quantities? -->
![ The flow of association and causation in graphs. Extracted from @Neal_2020 [pp. 31] ](images/figures/ACflow.png){#fig-ACflow fig-align="center" width=100%}

<!-- 4. What is the local markov assumption? and where does it lead? -->
<!-- 4.1 Bayesian factorization -->

<!-- 5. What is the minimality assumption? and where does it lead? -->

<!-- 6. What is the causal edges assumption? and where does it lead? -->

<!-- 7. putting it all together -->

<!-- 9. dependence and independence in building blocks -->

<!-- 9.1 Blocked path definition-->

<!-- 9.2 d-separation -->

<!-- This further implies that to understand the causal assumptions of a complex system, one can decompose the structure into these basic building blocks and analyze the causal assumptions associated with each block individually [@McElreath_2020]. -->

>>>>

<!-- 3. What are the advantages of DAGs? -->
DAGs offer two key advantages for modeling causality. Firstly, DAGs represent causal relations in a nonparametric and fully interactive manner. This feature implies that feasible causal analysis strategies usually do not require specifying the type of data or the nature of the functional dependence among variables [@Morgan_et_al_2014]. Secondly, regardless of complexity, DAGs can represent various causal structures using only five fundamental building blocks [@Neal_2020; @McElreath_2020]. This further implies that to understand the causal assumptions of a complex system, one can decompose a structure into these basic building blocks and analyze the causal assumptions associated with each block individually [@McElreath_2020]. Importantly, these building blocks can be represented in three ways: the magnified representation, the standard representation, and the structural causal model form [SCM, @Morgan_et_al_2014].

@@

<!-- 1. But what are the assumption that allow us to treat DAGs as causal structures -->
![ The flow of association and causation in graphs. Extracted from @Neal_2020 [pp. 31] ](images/figures/ACflow.png){#fig-ACflow fig-align="center" width=100%}

<!-- 2. What is the local markov assumption? and where does it lead? -->
<!-- 2.1 Bayesian network factorization -->

<!-- 3. What is the causal edges assumption? and where does it lead? -->

<!-- 4. Combined they are known as the minimality assumption? -->

<!-- 5. Now, what are those mysterious functions?, and how do they become statistical quantities? -->

<!-- 6. What is the modularity assumption? and where does it lead? -->
<!-- 6.1 the Truncated factorization -->
<!-- 6.2 The law of counterfactuals -->

<!-- 7. Causal mechanism -->
<!-- 7.3 the backdoor criterion (do you need it?) -->

<!-- 5. d-separation -->
<!-- 5.1 Blocked path definition-->

<!-- 6. putting it all together -->
<!-- 6.1. Revising the building block: dependence and independence -->
